\textbf{Are regime changes always bad economics? Evidence from daily financial data}

### 2018-12-29 doc/paper.tex
Update title of paper
Test overleaf based commit
Some updates to paper including the introduction, data, and results sections

\setcounter{tocdepth}{1}
\begin{document}
\author{Devin Incerti and Trevor Incerti}
\title{\textbf{Political Instability and Financial Markets}}
\date{\today}
\maketitle

@@

Political instability is also thought to harm investment. For investors, multinational firms, development agencies, and aid organizations, political risk is a top consideration when making investment decisions in emerging markets.\footnote{In the 2013 Multilateral Investment Guarantee Agency (MIGA) \citet{wipr2013} report (the last report published), executives of multinational enterprises (MNEs) ranked political risk as the second most important constraint for foreign direct investment (FDI) in developing countries over the next three years (after macroeconomic instability). In \citet{wipr2011} and  \citet{wipr2012}, political risk was ranked the most important constraint for FDI in developing countries - even greater than macroeconomic instability. The types of political risk of most concern to investors in developing countries (ranked in order of importance) were 1) adverse regulatory changes, 2) breaches of contract, 3) transfers and convertibility restrictions, 4) civil disturbances, 5) non-honoring of government guarantees, 6) expropriation nationalization, 7) terrorism and 8) war.} These investor sentiments are supported by empirical research that suggests political instability - as measured by regime change frequency or political violence - is negatively correlated with investment and financial development in cross country regressions \citep{alesina1996income,roe2011political}. In addition, the variance of stock market returns is also estimated to increase in response to political instability.\footnote{\citet{leblang2005government} show that right-wing governments increase the mean and volatility of stock prices in the United States and Britain, while \citet{jensen2005market} find that a rise in the popularity of a Brazilian presidential candidate with uncertain future policies increased the volatility of the Brazilian stock market.}

Political instability is also hypothesized to reduce economic growth by increasing policy uncertainty. This hypothesis is once again supported by cross country empirical evidence that political instability is negatively correlated with GDP growth \citep{alesina1996political, fosu1992political, aisen2013does, jong2009measurement}. For example, \citet{alesina1996political} show that countries have lower growth in years in which there are major government changes such as coups. The exception to this conclusion is \citet{londregan1990poverty}, who find that coups have no effect on economic growth. Additional research focuses on government policy instruments that affect private rates of return on investment, and argues that political instability causes governments to act myopically and adopt inefficient policies.\footnote{For instance, \citet{svensson1998investment} argues that governments are less likely to invest in the legal system and the protection of property rights when political uncertainty is high. Similarly, \citet{devereux1998political} propose a model in which political instability causes governments to leave fewer assets to their successors which forces them into increasing capital taxes. The knowledge of future taxation causes the private sector to reduce current investment which reduces future output.}

However, such hypotheses are often overly simplistic. Not all political instability is equivalent - resignations, assassinations, and coups are all examples of political instability, but may have disparate effects on financial returns. Resignations typically occur when an unpopular leader steps down and may signal the coming of a more effective leader. Assassinations are not always related to the effectiveness of a leader, can occur seemingly at random, and the successor to the assassinated leader may be unclear. Coups can occur in the name of democracy or autocracy, and in generally signal relative weakness in the current political system. Further, political instability may be seen as a positive event if the current regime is regarded as strongly anti-business or anti-global.  \citet{alesina1996political} take the possibility of a new government adopting better economic policies seriously, but argue that the negative effects of uncertainty dominate the positive effects of coups staged by pro growth factions.

\bigskip

We use financial performance as an indicator of whether the market views political stability as ``good'' for economic growth. As in \citet{alesina1996political}, we use ``irregular'' regime changes such as coups, assassinations and resignations as an indicator for political instability. The major innovation of this paper is that, in contrast to the aforementioned cross country studies, we analyze the impact of political instability using daily financial data. This enables us to use market expectations to quantify the economic impact of each individual type of regime change. The advantage of this event study approach is that it allows us to isolate the effect of irregular regime changes and mitigate the endogeneity problems that are nearly ubiquitous in cross country regressions.\footnote{A growing body of work uses event studies to assess political phenomenon.  For example, political events have been used to estimate the effect of political connections on firm value \citep[e.g.][]{fisman2001estimating,faccio2006politically,goldman2009politically}. Studies in ``forensic economics" have used abnormal returns from political events to locate or examine transactions such as insider trading \citep{dube2011coups}, illegal arms trading \citep{dellavigna2010detecting}, and the impact of hostilities on the financial performance of diamond mining firms in Angola \citep{guidolin2007diamonds}.}

@@

\section{Data}
Political data are primarily drawn from the Center for Systemic Peace's (CSP) Polity IV Coup d'etat dataset and Coup d'etat Events handbook. The Coup d'etat dataset includes the date of 1) successful coups, 2) attempted coups, 3) plotted coups and 4) alleged coup plots. We focus on successful coups because it is difficult to classify failed coups.\footnote{\citet[p. 617]{needler1966political} has even gone so far as to say that ``the categories of coups that were aborted, suppressed, or abandoned melt into each other and into a host of other non-coup phenomena so as to defy accounting.''} As even successful coups can have disparate outcomes (e.g. a shift from authoritarian rule to democratic rule, or vica versa), we provide a list of coups and their general outcome as defined by the Polity IV Project in \autoref{tab:coup list} below.

\singlespacing
\footnotesize
\begin{center}
\begin{longtable}[!ht]{l@{\extracolsep{\fill}}lll}
\caption{List of Coup d'Etat}
\label{tab:coup list}\\
\vspace{-5pt}\\
\hline
\multicolumn{1}{c}{Coup Date}&\multicolumn{1}{c}{Country}&\multicolumn{1}{c}{Coup Outcome}\\\hline
\endfirsthead
\multicolumn{3}{c}%
{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
\hline
\hline
\multicolumn{1}{c}{Coup Date}&\multicolumn{1}{c}{Country}&\multicolumn{1}{c}{Coup Outcome}\\\hline
\hline
\endhead
\hline 
\multicolumn{3}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot
06/08/1970 & Argentina & Autocracy to autocracy\\
03/22/1971 & Argentina & Autocracy to autocracy (democracy by 1973)\\
03/24/1976 & Argentina & Democracy to autocracy\\
10/06/1976 & Thailand & Anocracy to autocracy\\
10/20/1977 & Thailand & Autocracy to autocracy\\
12/12/1979 & South Korea & Autocracy to autocracy\\
02/23/1991 & Thailand & Anocracy to anocracy\\
04/05/1992 & Peru & Democracy to anocracy\\
10/12/1999 & Pakistan & Democracy to autocracy\\
10/04/2002 & Nepal & Democracy to autocracy\\
09/19/2006 & Thailand & Democracy to anocracy\\
01/11/2007 & Bangladesh & Democracy to autocracy\\
07/03/2013 & Egypt & Anocracy to anocracy\\
05/22/2014 & Thailand & Democracy to anocracy\\
 & \begin{tabular}[x]{@{}c@{}}\\\end{tabular} & \\
04/11/2002 & Venezuela & Partial/failed democracy to democracy\\
\hline
\hline
\end{longtable}
\end{center}
\doublespacing
\normalsize

The Coup d'etat Events handbook also provides a list of 1) auto-coups\footnote{Defined by Center for Systemic Peace as the ``occurrence of subversion of the constitutional order by a ruling (usually elected) executive and the imposition of an autocratic regime"}, 2) the ouster of leadership by foreign forces, 3) the ouster of leadership by rebel forces, 4) assassinations of the executive and 5) resignations of the executive due to poor performance and/or loss of authority. Daily financial data is available for countries in categories 4 and 5, so we supplement the coups with assassinations and resignations to form a list of ``irregular'' regime changes. The resignations are those in which the ruling executive was coerced to resign due to poor performance, public discontent and popular demonstrations.

@@

Financial data are from the Global Financial Data database, which includes the longest available time series of stock prices. We collect data on national equity indices and two global equity indices, the S\&P/IFC Emerging Market Investable Composite and the Morgan Stanley Capital International (MSCI) World Index. The S\&P/IFC index includes securities from emerging markets while the MSCI index includes securities from developed markets only. We collect stock index data on every country in which there was a coup or coup attempt, an assassination or failed assassination, or a forced resignation.\footnote{The list of failed assassinations are from \citet{jones2009hit}. Coup attempts are those in category 2 in the CSP Coup d'etat dataset.} The longest available daily time series for these stock indices are listed in \autoref{tab:stock list}.

\singlespacing
\footnotesize
\begin{center}
\begin{longtable}[!ht]{l@{\extracolsep{\fill}}lll}
\caption{List of stock indices}\label{tab:stock list}\\
\vspace{-5pt}\\
\hline
\hline
\multicolumn{1}{c}{Country}&\multicolumn{1}{c}{Stock Index}&\multicolumn{1}{c}{Begin Date}&\multicolumn{1}{c}{End Date}\\
\hline
\endfirsthead
\multicolumn{4}{c}%
{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
\hline
\hline
\multicolumn{1}{c}{Country}&\multicolumn{1}{c}{Stock Index}&\multicolumn{1}{c}{Begin Date}&\multicolumn{1}{c}{End Date}\\
\hline
\endhead
\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot
Argentina & Beunos Aires SE General Index & Dec-66 & Jan-17\\
Australia & Australia ASX All Ordinaries & Jan-58 & Jan-17\\
Bangladesh & Dhaka SE Index & Jan-90 & Jan-17\\
Canada & Canada S\&P/TSX 300 Composite & Jan-76 & Jan-17\\
Chile & Santiago IGBC General Index & Jan-75 & Jan-17\\
Colombia & Colombia IGBC General Index & Jan-92 & Jan-17\\
Ecuador & Ecuador Bolsa de Valores de Guayaquil (BVG) & Jan-94 & Jan-17\\
Egypt & Cairo SE Index & Dec-92 & Jan-17\\
Emerging Market & S\&P/IFC Emerging Markets Investable Composite & Jul-95 & Jan-17\\
Greece & Athens SE General Index & Oct-88 & Jan-17\\
India & Bombday SE SENSEX & Apr-79 & Jan-17\\
Indonesia & Jakarta SE Composite Index & Apr-83 & Jan-17\\
Iran & Tehran SE Price Index & Jan-95 & Jan-17\\
Israel & Tel Aviv 100 Index & May-87 & Jan-17\\
Japan & Tokyo SE Price Index (TOPIX) & Jan-53 & Jan-17\\
Latin America & Dow Jones Latin America Index & Jan-92 & Jan-17\\
Lithuania & Lithuania Lit-10 Index & Jan-99 & May-05\\
Malaysia & Malaysia KLSE Composite & Jan-80 & Jan-17\\
Nepal & Nepal NEPSE Stock Index & Jan-01 & Jan-17\\
Netherlands & Netherlands All-Share Price Index & Jan-80 & Jan-17\\
Pakistan & Karachi SE 100 Index & Jan-89 & Jan-17\\
Paraguay & Asuncion SE PDV General Index & Oct-93 & Sep-08\\
Peru & Lima SE General Index & Jan-82 & Jan-17\\
Philippines & Manila SE Composite Index & Jan-86 & Jan-17\\
Portugal & Oporto PSI-20 Index & Jan-86 & Jan-17\\
Singapore & Singapore FTSE ST Index & Jul-65 & Jan-17\\
South Korea & Korea SE Stock Price Index  & Jan-62 & Jan-17\\
Southeast Asia & Dow Jones Southeast Asia Index & Jan-92 & Jan-17\\
Spain & Madrid SE General Index & Aug-71 & Jan-17\\
Sri Lanka & Colombo SE All-Share Index & Dec-84 & Jan-17\\
Sweden & Sweden OMX Affarsvarlden General Index & Jan-80 & Jan-17\\
Taiwan & Taiwan SE Capitalization Weighted Index & Jan-67 & Jan-17\\
Thailand & Thailand SET General Index & Apr-75 & Jan-17\\
Tunisia & Tunisia SE Index & Dec-97 & Jan-17\\
Turkey & Instanbul IMKB 100 Price Index & Oct-87 & Jan-17\\
Ukraine & Ukraine PFTS OTC Index & Jan-98 & Jan-17\\
United Kingdom & UK FTSE All-Share Index & Nov-62 & Jan-17\\
United States & Dow Jones Industrial Average & Feb-1885 & Jan-17\\
Uruguay & Bolsa de Valores de Montevideo Index & Jan-08 & Jul-16\\
Venezuela & Dow Jones Venezuela Stock Index & Jan-92 & Jul-07\\
Venezuela & Caracas SE General Index & Jan-94 & Jan-17\\
World & MSCI World Price Index & Jan-76 & Jan-17\\
Zambia & Lusaka SE Index & Jan-02 & Apr-06\\
Zambia & Lusaka SE Index & Jul-11 & Jan-17\\

\hline
\hline
\end{longtable}
\end{center}
\doublespacing
\normalsize

@@

\section{Impact of Political Instability on Stock Returns} \label{sec: Impact of Political Instability on Stock Returns}
\subsection{Abnormal Returns} \label{subsec: Abnormal Returns}
We begin our analysis by studying the effect of irregular regime changes on stock returns. We follow the standard event study methodology as presented by, among others, \citet{mackinlay1997event} and \citet{campbell1997econometrics}. Normal performance is measured with a constant mean return model\footnote{A constant mean return model is used instead of a market model in order to maximize the number of observations (plausible market indices such as the MSCI World Index and the S\&P/IFC Emerging Markets Investable Composite Index only begin in 1976 and 1995 respectively). That said, results are insensitive to the use of a market model.},
\begin{align} \label{eqn:market-model}
R_{it}=\mu_{i}+\epsilon_{it},
\end{align}
where $R_{it}$ is the logged return of national stock index $i$ on trading day $t$ and $\epsilon_{it}$ is the error term. We calculate abnormal returns (ARs), in an ``event window'' surrounding the date of each coup, $AR_{i\tau}=R_{i\tau}-\widehat{\mu}_i$, where $\tau$ is a date in the event window, and $\widehat{\mu}_i$ is estimated in an ``estimation window'' preceding the event window with \autoref{eqn:market-model}. We use a 41 day event window (i.e. 20 pre-event trading days, the event day, and 20 post-event trading days). The estimation window is the 250 trading days prior to the start of the event window. The abnormal returns are then used to generate cumulative abnormal returns (CARs) between event day $\tau_1$ and event day $\tau_2$: $CAR(\tau_1,\tau_2)=\sum_{\tau=\tau_1}^{\tau_2}AR_{i\tau}$.

@@

\subsubsection{Coups} \label{subsec: Coups}
\autoref{tab:AR-coups} shows abnormal returns for national stock indices both preceding and following coup d'etat. \autoref{tab:AR-coups} contains all coups presented in \autoref{tab:coup list} with the exception of the Argentinian coup of March 24, 1976 and the partial/failed Venezuelan coup attempt of 2002. The March 24, 1976 Argentinian coup is excluded from our analysis because the stock market remained closed from March 24 to April 5, 1976, or a period of twelve days.\footnote{Treating this twelve day period as a single day CARs results in a positive abnormal return of 58\%, a fluctuation that seems qualitatively unreasonable.}  We later examine the Venezuelan coup attempt of 2002 in detail, but do not include it in the sample in \autoref{tab:AR-coups} as it does not match our definition of a successful coup.

@@

In contrast to coups and assassinations, abnormal returns following resignations are large and positive (see \autoref{tab:AR-resignations}). The mean event day abnormal return is over 4\% and event day abnormal returns are only negative and statistically significant at even the ten percent level in two out of the fifteen resignations (Pakistan on April 19, 1993 and Tunisia on June 31, 2011). However, these positive returns are short lived: the 7 day cumulative abnormal return is nearly identical to the event day abnormal return and the 20 day cumulative abnormal return is not statistically different from zero [[\textit{This doesn't look right to me. The positive returns seem persistent and to grow larger over time}]].

These results are also consistent with the idea that the expected impact of a political event on economic policy is highly important to the directionality of returns. The positive event day abnormal return following resignations is not surprising since resignations typically occurr because of poor performance and/or loss of authority. Leaders were often ousted following corruption charges, financial crises, and/or political violence.

@@

\subsection{Volatility} \label{subsec: Volatility}
Although irregular regime changes have disparate effects on the direction of stock returns depending on expected policy outcomes, the results from \autoref{subsec: Abnormal Returns} suggest that all of these political events increase financial volatility. However, since stock volatility is not directly observable, one must decide how to best estimate volatility. Our estimates are obtained from a generalized autoregressive conditional heteroskedasticity (GARCH) model estimated using 1000 pre-event days, the event day and 1000 post-event days. As in \citet{jensen2005market} and \citet{leblang2005government}, we use the GARCH (1,1) specification. In particular, for national stock index $i$,
\begin{align*}
R_{it}=\mu_i + \epsilon_{it},\hspace{1cm} \epsilon_{it}\sim \mathcal{N}\left(0,\sigma_{it}^2\right),
\end{align*}
where $\mu_i$ is a constant and,
\begin{align*}
\sigma_{it}^2&=\gamma_{i}+\alpha_{i}\epsilon_{i,t-1}^2+\beta_{i}\sigma_{i,t-1}^2.
\end{align*}
The key parameter of interest is the conditional variance, $\sigma_{it}^2$. The one-period-ahead volatility forecasts, $\sigma_{it}$, are larger when $\epsilon_{i,t-1}^2$ and $\sigma_{i,t-1}^2$ are larger. In other words, the model predicts that large shocks will be followed by other large shocks.

@@

\subsection{Robustness} \label{subsec: Robustness}
There are three potential concerns with the results in \autoref{subsec: Abnormal Returns}. First, the abnormal returns could have been driven by factors unrelated to the regime changes. Second, the reported means are based on small sample sizes so confidence intervals based on normally distributed abnormal returns may be inappropriate. Third, the true effects of irregular regime changes on firm value may be underestimated if investors had apriori information.

To explore the first concern, we create a synthetic control portfolio for each event based on the techniques introduced in \citet{abadie2003economic} and \citet{abadie2010synthetic}. Each country is given a weight which represents its influence in the synthetic control portfolio. The weight is chosen so that the daily returns and the variance of the daily returns of the control portfolio and the event country are most similar in the estimation window. The set of possible countries in the control portfolio consists of all countries listed in \autoref{tab:stock list}.\footnote{See the appendix for a more formal explanation.}

>>>>

\setcounter{tocdepth}{1}
\begin{document}
\author{Devin Incerti and Trevor Incerti}
\title{\textbf{Are regime changes always bad economics? Evidence from daily financial data}}
\date{\today}
\maketitle

@@

Of course, political instability may not only impact a country's political environment but its economy as well. For instance, investors, multinational firms, development agencies, and aid organizations, rank political risk as a top consideration when making investment decisions in emerging markets.\footnote{In the 2013 Multilateral Investment Guarantee Agency (MIGA) \citet{wipr2013} report (the last report published), executives of multinational enterprises (MNEs) ranked political risk as the second most important constraint for foreign direct investment (FDI) in developing countries over the next three years (after macroeconomic instability). In \citet{wipr2011} and  \citet{wipr2012}, political risk was ranked the most important constraint for FDI in developing countries - even greater than macroeconomic instability. The types of political risk of most concern to investors in developing countries (ranked in order of importance) were 1) adverse regulatory changes, 2) breaches of contract, 3) transfers and convertibility restrictions, 4) civil disturbances, 5) non-honoring of government guarantees, 6) expropriation nationalization, 7) terrorism and 8) war.} These investor sentiments are supported by empirical research that suggests political instability---as measured by regime change frequency or political violence---is negatively correlated with investment and financial development in cross country regressions \citep{alesina1996income,roe2011political} and that the variance of stock market returns increases in response to economic policy uncertainty \citep{leblang2005government, jensen2005market, liu2015economic}.

There is considerable evidence that political risk and policy uncertainty not only impact stock market returns, but economic growth as well. For example, cross country empirical evidence suggests that political instability is negatively correlated with GDP growth \citep{alesina1996political, fosu1992political, aisen2013does, jong2009measurement}, which may is likely because governments act myopically and adopt inefficient policies \citep{svensson1998investment,devereux1998political}.\footnote{\citet{svensson1998investment} argues that governments are less likely to invest in the legal system and the protection of property rights when political uncertainty is high. Similarly, \citet{devereux1998political} propose a model in which political instability causes governments to leave fewer assets to their successors which forces them into increasing capital taxes. The knowledge of future taxation then causes the private sector to reduce current investment which reduces future output.}

However, such hypotheses are often overly simplistic. Not all political instability is equivalent - resignations, assassinations, and coups are all examples of political instability, but may have disparate effects on financial returns. Resignations typically occur when an unpopular leader steps down and may signal the coming of a more effective leader. Assassinations are not always related to the effectiveness of a leader, can occur seemingly at random, and the successor to the assassinated leader may be unclear. Coups can occur in the name of democracy or autocracy, and in general, signal relative weakness in the current political system. Further, political instability may be seen as a positive event if the current regime is regarded as strongly anti-business or anti-global.  \citet{alesina1996political} take the possibility of a new government adopting better economic policies seriously, but argue that the negative effects of uncertainty dominate the positive effects of coups staged by pro growth factions. Likewise, \citet{londregan1990poverty}, find that coups have no impact on economic growth and hypothesize that this may because there is a bimodal distribution of coups, some of which enhance growth and others which restrict it.

In this study, examine the impact of ``irregular'' regime changes and public protests on economic performance and remain agnostic about the direction of the effect. In contrast to the aforementioned cross country studies, we  conduct our analysis using daily financial data. Market expectations in the form of stock market returns are therefore used as an indicator of whether the market views different types of potentially destabilizing political events as ``good'' for economic growth.  The advantage of this event study approach is that it allows us to use time-series variation in stock returns to estimate treatment effects and mitigate the endogeneity problems that are nearly ubiquitous in cross country regressions.\footnote{A growing body of work uses event studies to assess political phenomenon.  For example, political events have been used to estimate the effect of political connections on firm value \citep[e.g.][]{fisman2001estimating,faccio2006politically,goldman2009politically}. Studies in ``forensic economics" have used abnormal returns from political events to locate or examine transactions such as insider trading \citep{dube2011coups}, illegal arms trading \citep{dellavigna2010detecting}, and the impact of hostilities on the financial performance of diamond mining firms in Angola \citep{guidolin2007diamonds}.}

@@

\section{Data}
Political data are primarily drawn from the Center for Systemic Peace's (CSP) Polity IV Coup d'etat dataset and Coup d'etat Events handbook. The Coup d'etat dataset includes the date of 1) successful coups, 2) attempted coups, 3) plotted coups and 4) alleged coup plots. We focus on successful coups because it is difficult to classify failed coups.\footnote{\citet[p. 617]{needler1966political} has even gone so far as to say that ``the categories of coups that were aborted, suppressed, or abandoned melt into each other and into a host of other non-coup phenomena so as to defy accounting.''} The one exception is that we separately study the failed coup attempt in Venezuela in April 2002 in which the President of Venezuela, Hugo Chavez, was removed from office for two days because it provides a natural experiment that allows us to estimate the impact of the forceful removal of a left-wing populist with a pro-business regime.

\begin{table}[!ht]
\caption{Regime changes} \label{tab:regime-changes}
\vspace{-5pt}
\footnotesize
\begin{center}
\begin{threeparttable}
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ll}
  \hline
  \hline
  \multicolumn{1}{c}{Date}&\multicolumn{1}{c}{Country}&\multicolumn{1}{c}{Political Outcome}\\
  \hline
   \multicolumn{3}{c}{\textbf{Coups}}\\ 
\ExpandableInput{../tables/coups.txt} \\
   \multicolumn{3}{c}{\textbf{Failed coup}}\\ 
\ExpandableInput{../tables/failed_coup.txt}\\
  \multicolumn{3}{c}{\textbf{Assassinations}}\\ 
\ExpandableInput{../tables/assassinations.txt}\\
  \multicolumn{3}{c}{\textbf{Resignations}}\\ 
\ExpandableInput{../tables/resignations.txt}
   \hline
   \hline
\end{tabular*}
\scriptsize
Notes: The Polity score is used to classify political outcomes as follows: autocracy = $-10 \leq \rm{score} \leq -6$, anocracy = $-5 \leq \rm{score} \leq 5$, and democracy = $6 \leq \rm{score} \leq 10$.
\end{threeparttable}
\end{center}
\end{table}

The Coup d'etat Events handbook also provides a list of 1) auto-coups\footnote{Defined by Center for Systemic Peace as the ``occurrence of subversion of the constitutional order by a ruling (usually elected) executive and the imposition of an autocratic regime."}, 2) the ouster of leadership by foreign forces, 3) the ouster of leadership by rebel forces, 4) assassinations of the executive and 5) resignations of the executive due to poor performance and/or loss of authority. Daily financial data is available for countries in categories 4 and 5, so we supplement the coups with assassinations and resignations to form a list of ``irregular'' regime changes. The resignations are those in which the ruling executive was coerced to resign due to poor performance, public discontent and popular demonstrations.

@@

A list of the political events in our dataset is shown in \autoref{tab:regime-changes}. Coups tend to have the largest impact on the level of democratization as a number of countries have subeqeuently transitioned from democracies to anocracies or autocracies. On the other hand, neither assassinations or resignations have much of an impact on the level of democratization.

Financial data are from the Global Financial Data database, which includes the longest available time series of stock prices. We collect data on national equity indices and two global equity indices, the S\&P/IFC Emerging Market Investable Composite and the Morgan Stanley Capital International (MSCI) World Index. The S\&P/IFC index includes securities from emerging markets while the MSCI index includes securities from developed markets only. We collect stock index data on every country in which there was a coup or coup attempt, an assassination or failed assassination, or a forced resignation.\footnote{The list of failed assassinations are from \citet{jones2009hit}. Coup attempts are those in category 2 in the CSP Coup d'etat dataset.} The longest available daily time series for these stock indices are listed in \autoref{tab:stock-list}.

\begin{table}[!ht]
\caption{List of stock indices} \label{tab:stock-list}
\vspace{-5pt}
\footnotesize
\begin{center}
\begin{threeparttable}
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}lll}
  \hline
  \hline
  \multicolumn{1}{c}{Date}&\multicolumn{1}{c}{Country}&\multicolumn{1}{c}{Begin Date} &\multicolumn{1}{c}{Start Date}\\
  \hline
Argentina & Beunos Aires SE General Index & Dec-66 & Jan-17\\
Australia & Australia ASX All Ordinaries & Jan-58 & Jan-17\\
Bangladesh & Dhaka SE Index & Jan-90 & Jan-17\\
Canada & Canada S\&P/TSX 300 Composite & Jan-76 & Jan-17\\
Chile & Santiago IGBC General Index & Jan-75 & Jan-17\\
Colombia & Colombia IGBC General Index & Jan-92 & Jan-17\\
Ecuador & Ecuador Bolsa de Valores de Guayaquil (BVG) & Jan-94 & Jan-17\\
Egypt & Cairo SE Index & Dec-92 & Jan-17\\
Emerging Market & S\&P/IFC Emerging Markets Investable Composite & Jul-95 & Jan-17\\
Greece & Athens SE General Index & Oct-88 & Jan-17\\
India & Bombday SE SENSEX & Apr-79 & Jan-17\\
Indonesia & Jakarta SE Composite Index & Apr-83 & Jan-17\\
Iran & Tehran SE Price Index & Jan-95 & Jan-17\\
Israel & Tel Aviv 100 Index & May-87 & Jan-17\\
Japan & Tokyo SE Price Index (TOPIX) & Jan-53 & Jan-17\\
Latin America & Dow Jones Latin America Index & Jan-92 & Jan-17\\
Lithuania & Lithuania Lit-10 Index & Jan-99 & May-05\\
Malaysia & Malaysia KLSE Composite & Jan-80 & Jan-17\\
Nepal & Nepal NEPSE Stock Index & Jan-01 & Jan-17\\
Netherlands & Netherlands All-Share Price Index & Jan-80 & Jan-17\\
Pakistan & Karachi SE 100 Index & Jan-89 & Jan-17\\
Paraguay & Asuncion SE PDV General Index & Oct-93 & Sep-08\\
Peru & Lima SE General Index & Jan-82 & Jan-17\\
Philippines & Manila SE Composite Index & Jan-86 & Jan-17\\
Portugal & Oporto PSI-20 Index & Jan-86 & Jan-17\\
Singapore & Singapore FTSE ST Index & Jul-65 & Jan-17\\
South Korea & Korea SE Stock Price Index  & Jan-62 & Jan-17\\
Southeast Asia & Dow Jones Southeast Asia Index & Jan-92 & Jan-17\\
Spain & Madrid SE General Index & Aug-71 & Jan-17\\
Sri Lanka & Colombo SE All-Share Index & Dec-84 & Jan-17\\
Sweden & Sweden OMX Affarsvarlden General Index & Jan-80 & Jan-17\\
Taiwan & Taiwan SE Capitalization Weighted Index & Jan-67 & Jan-17\\
Thailand & Thailand SET General Index & Apr-75 & Jan-17\\
Tunisia & Tunisia SE Index & Dec-97 & Jan-17\\
Turkey & Instanbul IMKB 100 Price Index & Oct-87 & Jan-17\\
Ukraine & Ukraine PFTS OTC Index & Jan-98 & Jan-17\\
United Kingdom & UK FTSE All-Share Index & Nov-62 & Jan-17\\
United States & Dow Jones Industrial Average & Feb-1885 & Jan-17\\
Uruguay & Bolsa de Valores de Montevideo Index & Jan-08 & Jul-16\\
Venezuela & Dow Jones Venezuela Stock Index & Jan-92 & Jul-07\\
Venezuela & Caracas SE General Index & Jan-94 & Jan-17\\
World & MSCI World Price Index & Jan-76 & Jan-17\\
Zambia & Lusaka SE Index & Jan-02 & Apr-06\\
Zambia & Lusaka SE Index & Jul-11 & Jan-17\\
   \hline
   \hline
\end{tabular*}
\scriptsize
Notes: Standard errors are in parentheses. ``Days to rebound'' is the number of trading days following a negative stock return for the national stock index to return to pre-event level (it is calculated if the price decreases on the event day, not if the event day abnormal return is negative). Returns are inflation adjusted. 
\end{threeparttable}
\end{center}
\end{table}

@@

\section{Impact of Political Instability on Stock Returns} \label{sec: Impact of Political Instability on Stock Returns}
\subsection{Abnormal Returns} \label{subsec:abnormal-returns}
We begin our analysis by studying the effect of irregular regime changes on stock returns. We follow the standard event study methodology as presented by, among others, \citet{mackinlay1997event} and \citet{campbell1997econometrics}. Normal performance is measured with a constant mean return model\footnote{A constant mean return model is used instead of a market model in order to maximize the number of observations (plausible market indices such as the MSCI World Index and the S\&P/IFC Emerging Markets Investable Composite Index only begin in 1976 and 1995 respectively). That said, results are insensitive to the use of a market model.},
\begin{align} \label{eqn:market-model}
R_{it}=\mu_{i}+\epsilon_{it},
\end{align}
where $R_{it}$ is the logged return of national stock index $i$ on trading day $t$ and $\epsilon_{it}$ is the error term. We calculate abnormal returns (ARs), in an ``event window'' surrounding the date of each coup, $AR_{i\tau}=R_{i\tau}-\widehat{\mu}_i$, where $\tau$ is a date in the event window, and $\widehat{\mu}_i$ is estimated in an ``estimation window'' preceding the event window with \autoref{eqn:market-model}. We use a 41 day event window (i.e. 20 pre-event trading days, the event day, and 20 post-event trading days). The estimation window is the 250 trading days prior to the start of the event window. The abnormal returns are then used to generate cumulative abnormal returns (CARs) between event day $\tau_1$ and event day $\tau_2$: $CAR(\tau_1,\tau_2)=\sum_{\tau=\tau_1}^{\tau_2}AR_{i\tau}$.

@@

\subsubsection{Coups} \label{subsec: Coups}
\autoref{tab:AR-coups} shows abnormal returns for national stock indices both preceding and following coup d'etat. \autoref{tab:AR-coups} contains all coups presented in \autoref{tab:regime-changes} with the exception of the Argentinian coup of March 24, 1976 and the partial/failed Venezuelan coup attempt of 2002. The March 24, 1976 Argentinian coup is excluded from our analysis because the stock market remained closed from March 24 to April 5, 1976, or a period of twelve days.\footnote{Treating this twelve day period as a single day CARs results in a positive abnormal return of 58\%, a fluctuation that seems qualitatively unreasonable.}  We later examine the Venezuelan coup attempt of 2002 in detail, but do not include it in the sample in \autoref{tab:AR-coups} as it does not match our definition of a successful coup.

@@

In contrast to coups and assassinations, abnormal returns following resignations are large and positive (see \autoref{tab:AR-resignations}). The mean event day abnormal return is over 4\% and the positive returns are persistent and grow larger over time (mean 20-day CAR $\approx$ 11\%). Furthermore, event day ARs are only negative and statistically significant at even the ten percent level in two out of the fifteen resignations (Pakistan on April 19, 1993 and Tunisia on June 31, 2011).

These results are consistent with the idea that the expected impact of a political event on economic policy is highly important to the directionality of returns. The positive event day abnormal return following resignations is not surprising since resignations typically occur because of poor performance and/or loss of authority. Leaders were often ousted following corruption charges, financial crises, and/or political violence.

@@

\subsection{Volatility} \label{subsec: Volatility}
Although irregular regime changes have disparate effects on the direction of stock returns depending on expected policy outcomes, the results from \autoref{subsec:abnormal-returns} suggest that all of these political events increase financial volatility. However, since stock volatility is not directly observable, one must decide how to best estimate volatility. Our estimates are obtained from a generalized autoregressive conditional heteroskedasticity (GARCH) model estimated using 1000 pre-event days, the event day and 1000 post-event days. As in \citet{jensen2005market} and \citet{leblang2005government}, we use the GARCH (1,1) specification. In particular, for national stock index $i$,
\begin{align*}
R_{it}=\mu_i + \epsilon_{it},\hspace{1cm} \epsilon_{it}\sim \mathcal{N}\left(0,\sigma_{it}^2\right),
\end{align*}
where $\mu_i$ is a constant and,
\begin{align*}
\sigma_{it}^2&=\gamma_{i}+\alpha_{i}\epsilon_{i,t-1}^2+\beta_{i}\sigma_{i,t-1}^2.
\end{align*}
The key parameter of interest is the conditional variance, $\sigma_{it}^2$. The one-period-ahead volatility forecasts, $\sigma_{it}$, are larger when $\epsilon_{i,t-1}^2$ and $\sigma_{i,t-1}^2$ are larger. In other words, the model predicts that large shocks will be followed by other large shocks.

@@

\subsection{Robustness} \label{subsec: Robustness}
There are three potential concerns with the results in \autoref{subsec:abnormal-returns}. First, the abnormal returns could have been driven by factors unrelated to the regime changes. Second, the reported means are based on small sample sizes so confidence intervals based on normally distributed abnormal returns may be inappropriate. Third, the true effects of irregular regime changes on firm value may be underestimated if investors had apriori information.

To explore the first concern, we create a synthetic control portfolio for each event based on the techniques introduced in \citet{abadie2003economic} and \citet{abadie2010synthetic}. Each country is given a weight which represents its influence in the synthetic control portfolio. The weight is chosen so that the daily returns and the variance of the daily returns of the control portfolio and the event country are most similar in the estimation window. The set of possible countries in the control portfolio consists of all countries listed in \autoref{tab:stock-list}.\footnote{See the appendix for a more formal explanation.}
